How to be a good member of a scientific software community [Article v\versionnumber]

### 2021-08-19 paper.tex
more text -- at least something for every section, I think

This paper is primarily aimed at junior scientists who are new to performing
research and inexperienced at asking for help. However, we hope it will be
valuable to anyone who uses software to do their research and struggles to ask
for help.

@@

This paper has focused on what software consumers should do when asking for
help.  I suppose it could even be read as a list of complaints about bad
behavior by users (though that's not how I intend it). However, each behavior
described above has a corresponding expectation for the developers.

However, this does not mean that the developers of scientific
software are without responsibilities. If we want users to behave in a certain
way, it behooves us to tell them that. Users approaching us in good faith
deserve to be treated fairly and courteously.

connect to open science: FAIR principles say developers should want code to be
reusable. Publishing the code isn't enough, you need to make it usable as well.
That means support.

Treating everyone with courtesy and respect, regardless of their identity, is
essential if you want the software tool to have a vibrant community surrounding
it. Ensuring people aren't excluded is the fair and just way to proceed ---
making sure that racist, sexist, anti-LGBTQ, etc language and behavior is
unwelcome and unacceptable is simply the moral thing to do (it's also pragmatic
--- can you really afford to lose talented developers and users because they
don't feel welcome?) This doesn't just mean use of epithets in messages (though
that's clearly unacceptable). Keep an eye on how things are named, on jokes
present in the code or messages, on whose opinions are listened to and who is
ignored.  Above all, if someone tells you a particular behavior or situation is
harmful, \emph{believe them}. Something that feels minor (or is just a joke) to
you may hit others very differently, and part of respect is understanding and
accounting for that.

Moreover, it's not enough for you as a developer to treat people with respect.
To the extent that there's a community around your software, you must attempt to
make sure that community is respectful; as developer, your words and deeds carry
significant weight. If you call out bad behavior (or even marginal behavior),
that will go a long way towards setting a good tone for your community. If you
let it slide, chances are others will assume you implicitly endorse it.

>>>>

This paper is primarily aimed at junior scientists who are new to performing
computational research and inexperienced at asking for help. However, we hope it
will be valuable to anyone who uses software to do their research and struggles
to ask for help.

@@

This paper has focused on what software consumers should do when asking for
help.  I suppose it could even be read as a list of complaints about bad
behavior by users (though that's not how I intend it). However, each behavior
described above has a corresponding expectation for the developers. If we expect
the users to try to solve their own problem, it behooves us to write good
document, give examples, and (most importantly) keep them up to date. If we want
to be contacted via GitHub only, we need to make that information clear and easy
to find. If we are frustrated by bug reports that don't give us the information
we need, we should tell people what kinds of information we need (e.g. by
providing bug report templates). In order to receive proper credit, we need to
document the appropriate way to cite the software. All of this information needs
to be clearly written out and easy to find.

Doing these things is valuable from a practical perspective; making it easy for
people to interact with us efficiently, even if it costs us some effort up
front, will make those interactions more pleasant. Even setting that aside, it's
the right way to do science --- FAIR principles for software development
\cite{Lamprecht-2020} prioritize not just releasing code, but making it useable.
That means documenting it well and supporting people who try to use it.

Users approaching us in good faith deserve to be treated fairly and courteously.
Treating everyone with courtesy and respect, regardless of their identity, is
essential if you want to build a vibrant effective community.  If the community
is just you, the sole developer, then the solution is obvious --- when you
interact with users, do so in ways that make them feel valued and respected.

As the user and developer community grows, other people will get involved in
supporting users. It's still your obligation as a developer to maintain those
standards.  Making sure that racist, sexist, anti-LGBTQ, etc language and
behavior is unwelcome and unacceptable is simply the moral thing to do (it's
also pragmatic --- can you really afford to lose talented developers and users
because they don't feel welcome?) This doesn't just mean use of epithets in
messages (though that's clearly unacceptable). Keep an eye on how things are
named, on jokes present in the code or messages, on whose opinions are listened
to and who is ignored.  Above all, if someone tells you a particular behavior or
situation is harmful, \emph{believe them}.  Something that feels minor (or is
just a joke) to you may hit others very differently, and part of respect is
understanding and accounting for that.

Along the same lines, a developer who berates a user for asking a stupid
question, or is rude and dismissive in a technical discussion about a new
feature, is actively damaging the community, \emph{regardless of the technical
quality of their contributions.} It's not enough for you as a developer to treat
people with respect.  To the extent that there's a community around your
software, you must attempt to make sure that community is respectful; as
developer, your words and deeds carry significant weight. If you call out bad
behavior (or even marginal behavior), that will go a long way towards setting a
good tone for your community. If you let it slide, chances are others will
assume you implicitly endorse it.
